---
title: "A Report on Lake Ontario's Microbes"
author: "Misha Kazi"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

In this report, we will analyze how different water quality parameters affect the abundance of bacterial phyla.

# Set-up

## Loading packages

First, we'll load necessary packages:

```{r packages, warning=FALSE, message=FALSE}
# warning - doesn't display warning
# message - removes message chunk about all the packages loaded
# echo - runs code and outputs its output but hides the code in code chunks
# include - didn't include code or output, ran code chunk silently
# eval
library(tidyverse)
```

## Reading in DAta

Then, we'll read in our data. We produced this dataframe by joining the sample_data.csv and taxon_abundance.csv using an inner join.

```{r data}
sample_and_taxon <- read_csv("data/sample_and_taxon.csv")
```

# Analyzing the Data

Finally, we'll plot the relationship between temperature and overall microbial cell abundance.

```{r cells_vs_temp}
sample_and_taxon %>% 
  ggplot() +
  aes(x = temperature,
      y = cells_per_ml/1e6,
      color = env_group) +
  geom_point() +
  labs(x = "Temperature (C)",
       y = "Cells (million/ml)",
       title = "Are temperature and cell abundance linked?")
```

It appears that warmer waters promote higher microbial cell counts. Cell counts were especially high in the warms water of Septermber.

```{r data_summary, include=FALSE}
nSamples <- nrow(sample_and_taxon)

minTemp <- sample_and_taxon %>% 
  pull(temperature) %>% 
  min() %>% 
  round() ## round all values to nearest whole number

# Make an object called maxTemp which has the maximum temperature
maxTemp <- sample_and_taxon %>% 
  pull(temperature) %>% 
  max() %>% 
  round()

```

The above plot shows `r nSamples` samples. For these samples the minimum temp was `r minTemp`C and the maximum was `r maxTemp`C.

| Statistic         | Value        |
|-------------------|--------------|
| Number of Samples | `r nSamples` |
| Minimum Temp      | `r minTemp`  |
| Maximum Temp      | `r maxTemp`  |


```{r kable}
library(knitr)

sample_and_taxon %>% 
  filter(env_group == "Deep") %>% 
  select(cells_per_ml, temperature, Chloroflexi) %>% 
  kable()
```


Challenge:
Make a bullet point list
First line is your favorite coding language so far - in italics
Second line is your favorite function we've learned so far - bold
Third line is the next programming thing you want to learn - italics and bold


Switch to a numbered list
Find a markdown formatting tutorial and insert it as a hyperlink

## Bullet Point List:
- *Markdown*

- **join**

- ***SQL***


## Numbered List:
1. *Markdown*

2. **join**

3. ***SQL***

## Markdown Formatting Tutorial
[Markdown Formatting Tutorial](https://docs.github.com/en/get-started/writing-on-github/getting-started-with-writing-and-formatting-on-github/basic-writing-and-formatting-syntax)


